---
title: "config Package"
author: "Joe Bahr"
date: 2021-12-21
categories: ["R"]
tags: ["package-overview"]
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE)
```

### Package Overview

The `config` package creates an easy framework for managing hard coded values used throughout a workflow.  It also enables multiple configurations for your workflow, a potentially useful feature for development.

### Why is this package useful?

Hard coded values should be avoided at all costs.  We define a hard coded value as any instance of a value being entered directly into the code.  Even for quick projects, burying hard coded values within a script can lead to errors and future frustration.  

```{r, eval=FALSE}
library(tidyverse)

# example of using hard coded values
mtcars %>% 
  filter(mpg <= 25 &  ## AVOID
         cyl %in% c(4,6,8)) ## AVOID
```

Centralizing all of these values in a config file makes the inputs explicit and transparent to the original programmer and anyone else who inherits the code.  Values used multiple times throughout a script also only need to be changed in a single place, further reducing room for error.

### Setup

First, we must install and load the `config` package.

```{r setup, eval=FALSE}
install.packages("config")
```

```{r, message=FALSE}
library(config)
```

Before we can call any functions from the `config` package, we need to create a config file.  This file should be saved as "config.yml" in your project directory.  The structure of this file will resemble the following (can omit comments):

```{r config, eval = FALSE}
# each config file must have a default configuration
default:
  max_mpg: 25
  carmarker: "all"
  cyl_values: !expr c(4,6,8)

# here we have a configuration for our production runs  
prod:
  max_mpg: 25
  carmaker: "all"
  cyl_values: !expr c(4,6,8)

# and finally a configuration for our test runs
test:
  max_mpg: 20
  carmaker: "mercedes"
  cyl_values: !expr c(4)
```

The structure of the above config file is three separate configurations (default, prod, test).  Within each of these configurations we have key-value pairs explicitly defined.

The first two values are simply  single values for `max_mpg` and `carmaker`, with one being a numeric value and the other a string.  Yet for `cyl_values` we would like to include several values in our configuration, requiring a vector of multiple values.  

The `config` file allows us to write the actual R code to create this vector by prefacing the expression, `c(4,6,8)`, with `!expr`.  This particular value is simply a piece of R code with the results being assigned to `cyl_values` within the configuration.

Now, we will look at how we read in these values and work them into our scripts.

### Read Config Values

The `get()` function within the `config` package is able to read and parse a given config file.  For more information on the `get()` function, simply run `?config::get()` in your console.

This function has two key arguments we will use every time we call it:

  - **file**: filepath to the "config.yml" file
  - **config**: name of configuration to read, in this case we must choose either "default", "prod", or "test"
 
In this case, we are reading in the "prod" configuration.

```{r pressure, echo=TRUE}
# call config::get()
config <- config::get(file = "./config.yml", config = "prod")
```

Now, we can extract the value for max_mpg within the prod configuration.  By assigning it to the variable `max_mpg`, we can now use it throughout the remainder of our workflow.
```{r}
max_mpg <- config$max_mpg
print(max_mpg)
```

Same goes for `cyl_values`.
```{r}
cyl_values <- config$cyl_values
print(cyl_values)
```

### Using Values in a Workflow

Just like any other variable, we are free to use the values extracted from our config file within our workflow.

```{r, eval=FALSE}
mtcars %>% 
  filter(mpg <= max_mpg &
         cyl %in% cyl_values)

```

### Conclusion

Creating a config file is an effective way to gather all inputs in a single location and ultimately reduce room for error.  While it is also possible to simply gather and define all inputs at the top of a script, using a config file puts a bit more rigor behind it and enables multiple configurations.  I hope you find this tip useful and start incorporating a config file into your own workflows!